\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-05-13 doc/paper.tex
give more details on the fixed-point iteration

The $\beta$-functions for the model parameters together with the
boundary conditions for a boundary value problem.  The default
two-scale RG solver tries to find the full set of model parameters
consistent with all constraints at all scales.  It does so by running
iteratively between the scales of all boundary conditions, imposing
the constraints and checking for convergence after each iteration.
This approach is decribed in \cite{Barger:1993gh} for the MSSM and is
widely implemented in SUSY spectrum generators.

@@

follows, see \figref{fig:two-scale-algorithm}:

@@

\item At the $M_Z$ scale the SUSY model gauge couplings $g_{1,2,3}$
  are set to the known Standard Model values (ignoring threshold
  corrections).
\item Then, the user-defined low-scale boundary conditions (defined in
  \code{LowScaleInput}) are imposed.
\item The Yukawa couplings $y_{u,d,e}$ of the SUSY model are set to
  the known Standard Model Yukawa couplings.

@@

  \code{HighScaleInput}).

@@

  low-scale (\code{LowScale})

@@

  \item The low-scale is recalculated.
  \item The \DRbar\ gauge and Yukawa couplings $g_{1,2,3}(M_Z)$,
    $y_{u,d,e}(M_Z)$ of the SUSY model are calculated using threshold
    corrections.

@@

    (\code{LowScaleInput})

@@

\item Run all model parameters to the high-scale (\code{HighScale})

@@

  \item Recalculate the high-scale
  \item Apply the high-scale constraint (\code{HighScaleInput})

@@

\item Run model parameters to the SUSY-scale (\code{SUSYScale})

@@

  \item Calculate the \DRbar\ mass spectrum
  \item Recalculate the SUSY-scale
  \item Apply the SUSY-scale constraint (\code{SUSYScaleInput})
  \item Solve the EWSB eqs.\ at the loop level

@@

\item If not converged yet, goto \ref{rge-step-one}

@@

  \begin{tikzpicture}[node distance = 2cm, auto]

@@

    \path[arrow,dashed] (MS) -- node[text width=16em] {on convergence\\ run to \code{SUSYScale}} (spec);

>>>>

The RGEs for the model parameters together with the boundary
conditions form a boundary value problem.  The default two-scale RG
solver tries to find a set of model parameters consistent with all
constraints at all scales.  It does so by running iteratively between
the scales of all boundary conditions, imposing the constraints and
checking for convergence after each iteration.  This approach is
decribed in \cite{Barger:1993gh} for the MSSM and is widely
implemented in SUSY spectrum generators.

@@

follows, see also \figref{fig:two-scale-algorithm}:

@@

\item At the $M_Z$ scale the gauge couplings $g_{1,2,3}$ are set to
  the known Standard Model values (ignoring threshold corrections).
\item The user-defined initial guess at the low-scale (defined in
  \code{InitialGuessAtLowScale}) is imposed.  In the example given in
  \secref{Sec:modfile} the Higgs VEVs are set to
    v_d &= v \cos\beta, & v_u &= v \sin\beta ,
  where $v=246.22\unit{GeV}$.  Afterwards, the Yukawa couplings
  $y_{u,d,e}$ of the SUSY model are set to the known Standard Model
  Yukawa couplings (ignoring SUSY particle corrections).

@@

  \code{HighScaleInput}).  Afterwards, the user-defined initial guess
  for the remaining model parameters (defined in
  \code{InitialGuessAtHighScale}) is imposed.  In the example given in
  \secref{Sec:modfile} the superpotential parameter $\mu$ is set to
  the value $1.0$ and its corresponding soft-breaking parameter $B\mu$
  is set to zero.

@@

  low-scale (\code{LowScale}).

@@

  \item The low-scale is recalculated.  In the above example this step
    is trivial, because the low-scale is fixed to be $M_Z$.
  \item The \DRbar\ gauge couplings $g_{1,2,3}(M_Z)$ of the SUSY model
    are calculated using threshold corrections as described in
    \secref{sec:calculation-of-gauge-couplings}.

@@

    (\code{LowScaleInput}).  In the example above, the Yukawa
    couplings are calculated automatically as described in
    \secref{sec:calculation-of-yukawa-couplings} and the Higgs VEVs
    are set to
      v_d(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \cos\beta(M_Z)}, \\
      v_u(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \sin\beta(M_Z)}.
    Since the Hypercharge gauge coupling $g_1$ is GUT normalized, the
    normalization factor $\sqrt{3/5}$ has to be included in the above
    relations.

@@

\item Run all model parameters to the high-scale (\code{HighScale}).

@@

  \item Recalculate the high-scale as
      M_X' = M_X \exp\left(\frac{g_2(M_X)-g_1(M_X)}{\beta_{g_1} -
          \beta_{g_2}}\right),
    where $\beta_{g_i}$ is the two-loop $\beta$-function of the gauge
    coupling $g_i$.  The value $M_X'$ is used as new high-scale in the
    next iteration.
  \item Impose the high-scale constraint (\code{HighScaleInput}).  In
    the MSSM example the following soft-breaking parameters are fixed
    to the universal values $m_0$, $M_{1/2}$ and $A_0$:
      A^f(M_X) &= A_0 y_f(M_X) & &(f=u,d,e),\\
      m_{H_i}^2(M_X) &= m_0^2 & &(i=1,2),\\
      m_{f}^2(M_X) &= m_0^2\mathbf{1} & &(f=q,l,d,u,e),\\
      M_{i}(M_X) &= M_{1/2} & &(i=1,2,3).
    \end{align}\label{eq:fs-cmssm-high-scale-bc}%

@@

\item Run model parameters to the SUSY-scale (\code{SUSYScale}).

@@

  \item Calculate the \DRbar\ mass spectrum.
  \item Recalculate the SUSY-scale $M_S$ as
      M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}},
    where $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the lightest and
    heaviest up-type squarks, respectively.
  \item Impose the SUSY-scale constraint (\code{SUSYScaleInput}).  In
    the example above, this step is trivial since
    \code{SUSYScaleInput} is set to be empty.
  \item Solve the EWSB equations iteratively at the loop level.  In
    the MSSM example from above leading two-loop corrections have been
    enabled by setting \code{UseHiggs2LoopMSSM = True;}.  This will
    add two-loop tadpole contributions to the effective Higgs
    potential during the EWSB iteration.

@@

\item If not converged yet, goto \ref{rge-step-one}.  Otherwise,
  finish the iteration.

@@

  \begin{tikzpicture}[node distance = 2.2cm, auto]

@@

    \path[arrow,dashed] (MS) -- node[text width=16em] {if converged run to \code{SUSYScale}} (spec);
